---
title: "synthReact Package Demo"
author: "Matthew Whitaker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = "figures/",
  fig.cap = TRUE
)
```


# Overview

This document demonstrates the usage of the `synthReact` R package, which generates synthetic datasets based on the REACT-1 study parameters. The package allows users to create realistic synthetic data for analysis and modeling while preserving data privacy.

# Installation

```{r installation, eval=FALSE}
# Install devtools if not already installed
install.packages("devtools")

# Install synthReact from GitHub
devtools::install_github("mathzero/synthReact")
```


# Loading Necessary Packages and Functions

```{r load_packages}
# Load required packages
library(synthReact)
library(ggplot2)
library(dplyr)
library(pROC)
library(corrplot)
```
# Generating Synthetic Data

Generate a synthetic dataset with 1,000 samples:

```{r generate_data}
# Set seed for reproducibility
seed_value <- 123

# Generate synthetic data
n_samples <- 10000
synthetic_data <- generate_synthetic_data(n_samples = n_samples, seed = seed_value)

# View the first few rows
head(synthetic_data)
```

# Data Exploration and Visualization

## Histogram of Age

```{r histogram_age, fig.cap="Histogram of Age"}
library(ggplot2)

ggplot(synthetic_data, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(x = "Age", y = "Count") +
  theme_minimal()
```

## Boxplot of BMI by Gender

```{r boxplot_bmi_gender, fig.cap="BMI by Gender"}
ggplot(synthetic_data, aes(x = gender, y = bmi)) +
  geom_boxplot(fill = "orange", color = "black") +
  labs(x = "Gender", y = "BMI") +
  theme_minimal()
```

## Bar Plot of Smoking Status

```{r barplot_smoking_status, fig.cap="Smoking Status Distribution"}
ggplot(synthetic_data, aes(x = smokeever_cat)) +
  geom_bar(fill = "green", color = "black") +
  labs(x = "Smoking Status", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Correlation Matrix of Continuous Variables

```{r correlation_matrix, fig.cap="Correlation Matrix of Continuous Variables"}
library(corrplot)
library(dplyr)

# Select continuous variables
continuous_vars <- synthetic_data %>%
  select(age, bmi, imd_decile, num_comorbidities)

# Compute correlation matrix
cor_matrix <- cor(continuous_vars)

# Visualize the correlation matrix
corrplot(cor_matrix, method = "color", type = "upper",
         tl.col = "black", tl.srt = 45,
         addCoef.col = "black", number.cex = 0.7,
         mar = c(0, 0, 1, 0))
```

# Modeling Outcomes

## Logistic Regression Model for PCR Test Result

```{r model_pcr_test_result}

# Convert outcome to factor
# synthetic_data$pcr_test_result <- factor(synthetic_data$pcr_test_result, levels = c(0, 1))

# Fit the model
model_pcr_test_result <- glm(as.formula("pcr_test_result ~ age + symptom_loss_smell+symptom_loss_taste+
                               symptom_cough + symptom_fever + covid_before+vaccinated"),
                             data = synthetic_data, family = binomial)

# Summary of the model
summary(model_pcr_test_result)
```

## Odds Ratios and Confidence Intervals

```{r odds_ratios_pcr_test_result}
# Odds ratios and confidence intervals
exp(cbind(OddsRatio = coef(model_pcr_test_result), confint(model_pcr_test_result)))
```

## Logistic Regression Model for Long COVID

```{r model_longcovid}
# Convert outcome to factor
synthetic_data$longcovid <- factor(synthetic_data$longcovid, levels = c(0, 1))

# Fit the model
model_longcovid <- glm(longcovid ~ age + gender + bmi + smokeever_cat + num_comorbidities,
                       data = synthetic_data, family = binomial)

# Summary of the model
summary(model_longcovid)
```

## Odds Ratios and Confidence Intervals

```{r odds_ratios_longcovid}
# Odds ratios and confidence intervals
exp(cbind(OddsRatio = coef(model_longcovid), confint(model_longcovid)))
```

# ROC Curve Analysis

## Predicted Probabilities and ROC Curve for PCR Test Result

```{r roc_curve_pcr_test_result, fig.cap="ROC Curve for PCR Test Result"}
library(pROC)

# Predicted probabilities for PCR Test Result
synthetic_data$predicted_prob_pcr <- predict(model_pcr_test_result, type = "response")

# ROC Curve for PCR Test Result
roc_pcr_test <- roc(as.numeric(synthetic_data$pcr_test_result), synthetic_data$predicted_prob_pcr)

# Plot ROC Curve
plot(roc_pcr_test, col = "red", main = "ROC Curve for PCR Test Result")

```

## AUC Value

```{r auc_value_pcr_test_result}
# AUC Value
auc_value_pcr <- auc(roc_pcr_test)
print(paste("AUC for PCR Test Result:", round(auc_value_pcr, 3)))
```

# Conclusion

This demo showcases how to generate and analyze synthetic data using the `synthReact` package. The synthetic data preserves the statistical properties of the original REACT-1 study data, allowing for meaningful analysis while ensuring data privacy.

---

# Additional Notes

- **Plots**: The images included in this document are generated when you knit the R Markdown file. Ensure that you have the necessary packages installed.

- **Package Dependencies**: Make sure to install all required packages before running the script:

  ```{r install_packages, eval=FALSE}
  install.packages(c("ggplot2", "dplyr", "pROC", "corrplot"))
  ```

- **Data Privacy**: The synthetic data generated does not contain any real participant information from the REACT-1 study, maintaining confidentiality.

# References

- **synthReact GitHub Repository**: [synthReact](https://github.com/mathzero/synthReact)

- **REACT-1 Study**: [Official Website](https://www.imperial.ac.uk/medicine/research-and-impact/groups/react-study/studies/the-react-1-programme/)

---
